---
title: "rfm.Rmd"
author: "Kieran healy"
date: "11/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data

The RFM table is a PDF file. We need to extract it and clean it up. I converted it to a CSV in Excel.

```{r}
library(tidyverse)

rfm_tbl <- read_csv("data/rfm_table.csv")

rfm_tbl 
  
```

Clean up:

1. Convert column names to lower/snake case
2. Remove blank lines
3. Separate the r, f, and m columns into pairs with their lower bound (if any) and upper bound. Names are e.g. `r_lo`, `r_hi` etc. The `separate()` function is very handy for splitting weirdly formatted columns into tractable data.
4. Move the original columns to the end.
5. Convert missing (`NA`) to zero: I'm assuming here that this is valid and that R, F, and M scores cannot be negative! 

```{r}
rfm_tbl <- rfm_tbl %>%
  janitor::clean_names() %>%
  filter_all(any_vars(!is.na(.))) %>%
  separate(col = r, into = c("r_lo", "r_hi"), 
           remove = FALSE, convert = TRUE, 
           fill = "left") %>%
  separate(col = f, into = c("f_lo", "f_hi"), 
           remove = FALSE, convert = TRUE, 
           fill = "left") %>%
  separate(col = m, into = c("m_lo", "m_hi"), 
           remove = FALSE, convert = TRUE, 
           fill = "left") %>%
  mutate_if(is.integer, replace_na, 0) %>%
  select(segment, r_lo, r_hi, f_lo, f_hi, m_lo, m_hi, everything())
  
rfm_tbl

```

Much nicer now. 

Looking at the criteria, this seems like a weird classification scheme! 

For now, I won't use these directly in the function below, but one could.

# Generate some Customer Score Data

```{r}

vals <- c(0:5)
n = 1000

customers <- tibble(r = sample(vals, n, replace = TRUE),
                    f = sample(vals, n, replace = TRUE),
                    m = sample(vals, n, replace = TRUE)
                    ) 

customers
```

## Classify triplets of r, f, and m scores

```{r}

calc_segment <- function(dat = customers){
  case_when(
    dat$r >= 4 & dat$r <= 5 & dat$f >= 4 & dat$f <= 5 & dat$m >= 4 & dat$m <= 5 ~ "Champion",
    dat$r >= 2 & dat$r <= 5 & dat$f >= 3 & dat$f <= 5 & dat$m >= 3 & dat$m <= 5 ~ "Loyal Customer",
    dat$r >= 3 & dat$r <= 5 & dat$f >= 1 & dat$f <= 3 & dat$m >= 1 & dat$m <= 3 ~ "Potential Loyalist",
    dat$r >= 4 & dat$r <= 5 & dat$f >= 0 & dat$f <= 1 & dat$m >= 0 & dat$m <= 1 ~ "New Customer",
    dat$r >= 3 & dat$r <= 4 & dat$f >= 0 & dat$f <= 1 & dat$m >= 0 & dat$m <= 1 ~ "Promising",
    dat$r >= 2 & dat$r <= 3 & dat$f >= 2 & dat$f <= 3 & dat$m >= 2 & dat$m <= 3 ~ "Need Attention",
    dat$r >= 3 & dat$r <= 3 & dat$f >= 0 & dat$f <= 2 & dat$m >= 0 & dat$m <= 2 ~ "About to Sleep",
    dat$r >= 0 & dat$r <= 2 & dat$f >= 2 & dat$f <= 5 & dat$m >= 2 & dat$m <= 5 ~ "At Risk",
    dat$r >= 0 & dat$r <= 1 & dat$f >= 4 & dat$f <= 5 & dat$m >= 4 & dat$m <= 5 ~ "Can't Lose Them",
    dat$r >= 1 & dat$r <= 2 & dat$f >= 1 & dat$f <= 2 & dat$m >= 1 & dat$m <= 2 ~ "Hibernating",
    dat$r >= 0 & dat$r <= 2 & dat$f >= 0 & dat$f <= 2 & dat$m >= 0 & dat$m <= 2 ~ "Lost",
    TRUE ~ "Other"
  )  
}

```

## Apply it to the data

```{r}
customers <- customers %>% 
  add_column(segment = calc_segment(customers))
  
customers
```

```{r}
customers %>% 
  group_by(segment) %>% 
  tally() %>%
  arrange(desc(n))
```

I wrote the criteria quickly so I'm not sure if this is right. But in general `case_when()` is a useful function to look at when you have multiple `if_else()` statements that you want to test a bunch of rows on.



